Document knowledge base linking for information retrieval

### 2016-02-12 paper.tex
minor updates

For the purposes of ad hoc document retrieval, however, the granularity of linking entities to specific mention spans in the text is not necessarily required. Since retrieval is concerned with scoring documents in their entirety, we may simplify the task from mapping \textit{mentions} to entities, as is more common, to mapping \textit{documents} to entities. We argue that such a ``bag of links'' provides many of the same benefits as more fine-grained entity linking or wikification processes for the purposes of ad hoc document retrieval while improving computational efficiency.

@@

Entity linking and wikification have been studied extensively, particularly in the context of TAC KBP \cite{Ji2014}. Entity linking systems often exploit knowledge base structure to help match and disambiguate entities. For example, Cucerzan \cite{Cucerzan2007} employs Wikipedia disambiguation pages and redirection pages to help identify various entity surface forms. The use of anchor text and hyperlink graphs is also a frequent feature of these systems \cite{Mihalcea2007, Ratinov2011, Stoyanov2012}. Most systems employ some form of context, which often refers to the co-occurrence of entities as evidence for disambiguation \cite{Stoyanov2012, Dalton2013}. Another use of context, relating most closely to our approach in this paper, refers the text surrounding an entity mention, which can be used to disambiguate knowledge base entries \cite{Mihalcea2007, Dalton2013}.

@@

Prior applications of knowledge base links in information retrieval use entity links in a feedback process for query expansion. Xiong and Croft, for example, employ FACC1\footnote{http://lemurproject.org/clueweb09/FACC1/} annotations on the document collection with supervised learning to expand queries \cite{Xiong2015}. Other research has used entity links in the queries themselves, rather than documents, to aid in expansion \cite{Xu2009, Brandao2014}. Still others use entity links in \textit{both} the query and documents \cite{Dalton2014, Li2015} to expand queries. Apart from entity linking, knowledge bases have also been used for query expansion in several related tasks, such as document filtering \cite{Xie2015} and blog search \cite{Elsas2008, Weerkamp2009}.

@@

\subsection{Linking with Document Pseudo-Queries}\label{section.linking.queries}
%The KL-divergence retrieval model measures the distance between the query language model, $\theta_{D_Q}$ here, and the document language model, $\theta_E$, as follows:
%
%\begin{flalign*}
%	score(E,D_Q) 	&= -KL(\theta_{D_Q}||\theta_E) \\
%					&\rankeq \sum_{w \in V} P(w|\theta_{D_Q}) \log P(w|\theta_E)
%&\end{flalign*}
%
To find candidate entities to ``link" to a given document $D$, we begin by treating the text of $D$ as a pseudo-query which we pose against a collection of entities $C_E$.  To transform a document into a pseudo-query we apply two transformations.  First we remove all terms from $D$ that appear in the standard Indri stoplist\footnote{http://www.lemurpro ject.org/stopwords/stoplist.dft}.  Next, we prune our pseudo-query by retaining only the $0 < k \leq |D|$ most frequent words in the stemmed text of $D$.  The integer variable $k$ is a parameter that we choose empirically.  Let $Q_D$ be the pseudo-query for $D$, consisting of the text of $D$ after our two transformations.

With our pseudo-query for document $D$ in hand, we obtain a list of candidate entities by running $Q_D$ over an index of our knowledge base, $C_E$, where each entry in this index is the text of an entity $E$'s knowledge base node. More formally, we rank the entities in our knowledge base against $D$ using Eq. \ref{equation.ql}, substituting $D_Q$ for the query and $E_i$--the text of the $i^{th}$ entity--for the document. Let $\pi_i$ be the log-probability for entity $E_i$ with respect to $D$ given by equation \ref{equation.ql}.

@@

\noindent Like $P(q_i|D)$, we estimate $P(q_i|E)$ as a Dirichlet-smoothed query likelihood, but using the knowledge base's text for entry $E$ to estimate of the underlying model. By virtue of our entity-document scoring and normalization, we also have $P(E|D)$.

@@

For comparison, we also report the results of our model using entity links produced by Apache Stanbol\footnote{http://stanbol.apache.org/}. We process the test collections with Stanbol's default ``enhancer,'' which supplies entity links and corresponding confidence scores. We select the link with the highest confidence for each mention. These links form the set of document links, and their confidence scores may be normalized to provide an estimate of $P(E|D)$.

@@

For the \textit{kb} and \textit{stanbol} runs, we retrieve the top 1000 documents per query based on the default Indri query likelihood implementation. We then re-rank these documents by incorporating their knowledge base links as described in Section \ref{section.model}.

@@

The various parameters required for our approach, along with their meanings and values are shown in Table \ref{table.parameters}.

@@

Retrieval performance of the top-scoring knowledge base runs and baselines are shown in Table \ref{table.performance}. Bolded mean average precision (MAP) scores are greater than the baseline run with statistical significance at $p < 0.001$ using a paired t-test. Note that baselines correspond to $\lambda = 0.0$.

Though Table \ref{table.performance} shows only the top-scoring values of $\lambda$, performance for \textit{kb} runs improved over the baseline for all values of $0.0 < \lambda < 1.0$, as shown in Figure \ref{figure.sweeps-ql}(a). All improvements were statistically significant with at least $p < 0.05$. We may explain the sudden drop in MAP at $\lambda=1.0$ by imagining a scenario in which two documents link to the same set of entities though only one of the documents is a good match for the query. While the entities provide the best data for query likelihood estimation, the loss of this document-specific quantity may be the cause of this drop. Further investigation is needed on this subject.

@@

The results indicate that our approach for constructing knowledge base links between documents and Wikipedia produces useful data for document retrieval purposes. Our simple document expansion model that incorporates these links performs well compared to a query likelihood baseline. These outcomes support our argument that a ``bag of links'' to a knowledge base can provide helpful information for a document retrieval task. Further, the poor performance of our model using more traditional mention-to-entity links indicates that, not only are document-to-entity links more efficient to produce, they also connect more useful knowledge base entries for the purposes of document retrieval.

In this paper, we have limited ourselves to using only knowledge base description text. However, knowledge base links provide a great deal more information. Future work may benefit from harnessing this knowledge base data such as hyperlink graphs and entity categories. Since our retrieval model performs \textit{document} expansion, we also plan to investigate its utility when paired with \textit{query} expansion techniques that employ knowledge base links.

>>>>

For the purposes of ad hoc document retrieval, however, the granularity of linking entities to specific mention spans in the text is not necessarily required. Since retrieval is concerned with scoring documents in their entirety, we may simplify the task from mapping \textit{mentions} to entities, as is more common, to mapping \textit{documents} to entities. We argue that such a ``bag of links'' provides many of the same benefits as more fine-grained entity linking or wikification processes for the purposes of ad hoc document retrieval while improving computational efficiency and robustness against overfitting.

@@

Entity linking and wikification have been studied extensively, particularly in the context of TAC KBP \cite{Ji2014}. Entity linking systems often exploit knowledge base structure to help match and disambiguate entities. For example, Cucerzan \cite{Cucerzan2007} employs Wikipedia disambiguation pages and redirection pages to help identify various entity surface forms. Most systems employ some form of context, which often refers to the co-occurrence of entities as evidence for disambiguation \cite{Stoyanov2012, Dalton2013}. Alternatively, context may refer to the text surrounding an entity mention, which can be used to disambiguate knowledge base entries \cite{Mihalcea2007, Dalton2013}.

@@

Prior applications of knowledge base links in information retrieval use entity links in a feedback process for query expansion. Xiong and Croft, for example, employ FACC1\footnote{http://lemurproject.org/clueweb09/FACC1/} annotations on the document collection with supervised learning to expand queries \cite{Xiong2015}. Other researchers use entity links in the queries themselves, rather than documents, to aid in expansion \cite{Xu2009, Brandao2014}. Still others use entity links in \textit{both} the query and documents \cite{Dalton2014, Li2015} to expand queries. Apart from entity linking, knowledge bases have also been used for query expansion in several related tasks, such as document filtering \cite{Xie2015} and blog search \cite{Elsas2008, Weerkamp2009}.

@@

\subsection{Linking with Document Pseudo-Queries}\label{section.linking.queries}
%The KL-divergence retrieval model measures the distance between the query language model, $\theta_{D_Q}$ here, and the document language model, $\theta_E$, as follows:
%
%\begin{flalign*}
%	score(E,D_Q) 	&= -KL(\theta_{D_Q}||\theta_E) \\
%					&\rankeq \sum_{w \in V} P(w|\theta_{D_Q}) \log P(w|\theta_E)
%&\end{flalign*}
%
To find candidate entities to ``link" to a given document $D$, we begin by treating the text of $D$ as a pseudo-query which we pose against a collection of entities $C_E$.  To transform a document into a pseudo-query we apply two transformations.  First we remove all terms from $D$ that appear in the standard Indri stoplist\footnote{http://www.lemurpro ject.org/stopwords/stoplist.dft}.  Next, we prune our pseudo-query by retaining only the $0 < k \leq |D|$ most frequent words in the stopped text of $D$.  The integer variable $k$ is a parameter that we choose empirically.  Let $Q_D$ be the pseudo-query for $D$, consisting of the text of $D$ after our two transformations.

We obtain a list of candidate entities by running $Q_D$ over an index of our knowledge base, $C_E$, where each entry in this index is the text of an entity $E$'s knowledge base node. More formally, we rank the entities in our knowledge base against $D$ using Eq. \ref{equation.ql}, substituting $Q_D$ for the query and $E_i$---the text of the $i^{th}$ entity---for the document. Let $\pi_i$ be the log-probability for entity $E_i$ with respect to $D$ given by Eq. \ref{equation.ql}.

@@

\noindent Like $P(q_i|D)$, we estimate $P(q_i|E)$ as a Dirichlet-smoothed query likelihood, but using the description text for entry $E$ to estimate of the underlying model. By virtue of our entity-document scoring and normalization, we also have $P(E|D)$.

@@

For comparison, we also report the results of our model using entity links produced by Apache Stanbol\footnote{http://stanbol.apache.org/}. We process the test collections with Stanbol's default ``enhancer,'' which supplies entity links and corresponding confidence scores. Though Stanbol links to DBpedia\footnote{http://dbpedia.org} resources, these generally correspond to Wikipedia pages and may be converted easily. We select the link with the highest confidence for each mention in a given document. These links form the set of document links, and their confidence scores may be normalized to provide an estimate of $P(E|D)$.

@@

We remove stop words in documents and entity descriptions for all runs. For the \textit{kb} and \textit{stanbol} runs, we retrieve the top 1000 documents per query based on the default Indri query likelihood implementation. We then re-rank these documents by incorporating their knowledge base links as described in Section \ref{section.model}.

@@

The various parameters required for our approach, along with their meanings and the values used in our experiments, are shown in Table \ref{table.parameters}.

@@

Retrieval performance of the baselines and top-scoring runs are shown in Table \ref{table.performance}. Bolded mean average precision (MAP) scores are greater than the baseline run with statistical significance at $p < 0.001$ using a paired t-test. Note that baselines correspond to $\lambda = 0.0$.

Though Table \ref{table.performance} shows only the top-scoring values of $\lambda$, performance for \textit{kb} runs improved over the baseline for all values of $0.0 < \lambda < 1.0$, as shown in Figure \ref{figure.sweeps-ql}(a). All improvements were statistically significant with at least $p < 0.05$. We may explain the sudden drop in MAP at $\lambda=1.0$ by imagining a scenario in which two documents link to the same set of entities though only one of the documents is a good match for the query. The loss of this direct document-query score may be the cause of the drop at $\lambda=1.0$ since, in this case, documents are scored only using their linked entities. Further investigation is needed on this subject.

@@

The results indicate that our approach for constructing knowledge base links between documents and Wikipedia produces useful data for document retrieval purposes. Our simple document expansion model that incorporates these links performs well compared to a query likelihood baseline. These outcomes support our argument that a ``bag of links'' to a knowledge base can provide helpful information for a document retrieval task. Further, the poor performance of our model using more traditional mention-to-entity links indicates that not only are document-to-entity links more efficient to produce, they also connect more useful knowledge base entries for the purposes of document retrieval.

In this paper, we have limited ourselves to using only knowledge base description text. However, knowledge base links provide a great deal more information. Future work may benefit from harnessing this knowledge base data, including hyperlink graphs and entity categories. Since our retrieval model performs \textit{document} expansion, we also plan to investigate its utility when paired with \textit{query} expansion techniques that employ knowledge base links.
